Computation and comparison of the stable Northeastern US marine boundary layer

### 2020-12-04 Paper/main.tex
Added nomenclature and fixed notationi

\section{Nomenclature}
{\renewcommand\arraystretch{1.0}
  \noindent\begin{longtable*}{@{}l @{\quad=\quad} l@{}}
ABL       & Atmospheric Boundary Layer \\
$\alpha$  & Wind shear exponent  \\
$f$       & Frequency \\
$f_{max}$  & Maximum resolvable frequency due to mesh limits \\
$f_{Ny}$   & Nyquist frequency \\
$L$       & Turbulent integral length scale \\
$L_{Ob}$   & Obukhov length scale \\
LES       & Large Eddy Simulation \\
$u_\tau$   & Friction velocity    \\
$\rho$    & Density \\
$\rho_\circ$ & Reference density  \\
$R_{ij}$   & Two point correlation tensor \\
SBL       & Stable Boundary Layer \\
$\theta$  & Potential temperature \\
$TI$      & Turbulence intensity \\
$TKE$     & Turbulent kinetic energy \\
$\overline{U}_{horiz}$ & Time averaged horizontal velocity
\end{longtable*}}

@@

In addition to these simulations, we also examined the domain and mesh
requirements necessary for the stable offshore boundary layers
considered here.  Based on the calculated turbulent integral
lengthscale, horizontal domain sizes of 750 m $\times$ 750 m were found
to be necessary for these simulations.  Various mesh resolutions were
also considered for the stable runs, and it was determined that the
finest resolution of $dx$=2.5 m was able to resolve the longitudinal
length scales and highest temporal frequencies, but inadequate for
resolving the turbulence in the lateral direction.  Further study is
warranted to comprehensively determine the impact of stable
stratification on the required mesh resolution.

>>>>

\section{Nomenclature}
{\renewcommand\arraystretch{1.0}
  \noindent\begin{longtable*}{@{}l @{\quad=\quad} l@{}}
ABL       & Atmospheric Boundary Layer \\
$\alpha$  & Wind shear exponent  \\
$\beta$   & Thermal expansion coefficient \\
$c_p$     & Specific heat capacity \\
$\Delta, \Delta_{x,y,z}$  & Grid spacing, grid spacing in x, y, z directions \\
$f$       & Frequency \\
$f_{max}$  & Maximum resolvable frequency due to mesh limits \\
$f_{Ny}$   & Nyquist frequency \\
$g, g_i$   & Gravitational acceleration \\
$k, TKE$  & Turbulent kinetic energy \\
$L$       & Turbulent integral length scale \\
$L_{Ob}$   & Obukhov length scale \\
$\nu_t$   & Turbulent viscosity \\
LES       & Large Eddy Simulation \\
$\rho$    & Density \\
$\rho_\circ$ & Reference density  \\
$\phi_h$  & Potential temperature gradient \\
$Pr_t$    & Turbulent Prandtl number \\
$R$       & Universal gas constant \\
$R_{ij}$   & Two point correlation tensor \\
SBL       & Stable Boundary Layer \\
$\theta$  & Potential temperature \\
$S_i$     & Wind spectra \\
$\sigma_i$ & Wind speed variance \\
$\tau_{ij}$ & Shear stress \\
$TI$      & Turbulence intensity \\
$\overline{U}_{horiz}$ & Time averaged horizontal velocity \\
$u_\tau$   & Friction velocity    \\
$V_1$     & Resolved wind speed at first grid level 
\end{longtable*}}

@@

In addition to these simulations, we also examined the domain and mesh
requirements necessary for the stable offshore boundary layers
considered here.  Based on the calculated turbulent integral
lengthscale, horizontal domain sizes of 750 m $\times$ 750 m were found
to be necessary for these simulations.  Various mesh resolutions were
also considered for the stable runs, and it was determined that the
finest resolution of $\Delta_x$=2.5 m was able to resolve the longitudinal
length scales and highest temporal frequencies, but inadequate for
resolving the turbulence in the lateral direction.  Further study is
warranted to comprehensively determine the impact of stable
stratification on the required mesh resolution.

### 2020-12-04 Paper/methodology.tex
Added nomenclature and fixed notationi

\subsection{Measured offshore conditions}
% \fxnote{Lawrence write this part}

@@

\subsubsection{\label{sec:wallmodelBC}Lower wall model BC}
The sub-filter scale stresses are specified at the terrain boundary following
Moeng~\cite{Moeng1984} as
\begin{align}
 (\tau_{xz})_0 &= \average{\tau_{xz}}_0 \frac{S_1 \average{\widetilde{u}_1} + \average{S_1} (\widetilde{u}_1 - \average{\widetilde{u}_1} ) }{ \average{S_1} \sqrt{\average{\widetilde{u}_1}^2 + \average{\widetilde{v}_1}^2 } } , \\
 (\tau_{yz})_0 &= \average{\tau_{yz}}_0 \frac{S_1 \average{\widetilde{v}_1} + \average{S_1} (\widetilde{v}_1 - \average{\widetilde{v}_1} ) }{ \average{S_1} \sqrt{\average{\widetilde{u}_1}^2 + \average{\widetilde{v}_1}^2 } } \\
  (\tau_{\theta z})_0 &= \average{\tau_{\theta z}}_0 \frac{S_1 ( \average{\widetilde{\theta}_1} - \theta_{\circ} ) + \average{S_1} (\widetilde{\theta}_1 - \average{\widetilde{\theta}_1} ) }{ \average{S_1} ( \average{\widetilde{\theta}_1} - \theta_{\circ} ) },
\end{align}
where all quantities with the subscript 1 are values at the first vertical grid
level, $S_1$ is the resolved wind speed at the first grid level and
$\average{\bullet}$ represents the horizontal planar average. The mean
sub-filter stresses are computed using Monin-Obukhov similarity
law~\cite{Etling1996}. The formulation for the potential temperature flux could
potentially experience a singularity when the mean resolved potential
temperature at the first grid level $\average{\widetilde{\theta}_1}$ crosses the
reference temperature $\theta_{\circ}$. Hence the model for the sub-filter
potential temperature flux is reformulated using the Monin-Obukhov similarity
law as
\begin{equation}
(\tau_{\theta z})_0 = \left ( \frac{u_{\tau} \kappa}{ \phi_h} \right ) \frac{S_1 ( \average{\widetilde{\theta}_1} - \theta_{\circ} ) + \average{S_1} (\widetilde{\theta}_1 - \average{\widetilde{\theta}_1} ) }{ \average{S_1} },
\end{equation}
where $\phi_h$ is the non-dimensional potential temperature gradient from
Monin-Obukhov similarity law~\cite{Etling1996}.

@@

\fxnote{Clarify we don't actually use adaptive refinement -- that is future work.}

@@

\subsection{\label{sec:CFDsetup}Computational setup}
% \fxnote{Domains, grids, boundary conditions}

>>>>

\subsection{Measured offshore conditions}

@@

\subsubsection{\label{sec:wallmodelBC}Lower wall model BC}
The sub-filter scale stresses are specified at the terrain boundary following
Moeng~\cite{Moeng1984} as
\begin{align}
 (\tau_{xz})_0 &= \average{\tau_{xz}}_0 \frac{V_1 \average{\widetilde{u}_1} + \average{V_1} (\widetilde{u}_1 - \average{\widetilde{u}_1} ) }{ \average{V_1} \sqrt{\average{\widetilde{u}_1}^2 + \average{\widetilde{v}_1}^2 } } , \\
 (\tau_{yz})_0 &= \average{\tau_{yz}}_0 \frac{V_1 \average{\widetilde{v}_1} + \average{V_1} (\widetilde{v}_1 - \average{\widetilde{v}_1} ) }{ \average{V_1} \sqrt{\average{\widetilde{u}_1}^2 + \average{\widetilde{v}_1}^2 } } \\
  (\tau_{\theta z})_0 &= \average{\tau_{\theta z}}_0 \frac{V_1 ( \average{\widetilde{\theta}_1} - \theta_{\circ} ) + \average{V_1} (\widetilde{\theta}_1 - \average{\widetilde{\theta}_1} ) }{ \average{V_1} ( \average{\widetilde{\theta}_1} - \theta_{\circ} ) },
\end{align}
where all quantities with the subscript 1 are values at the first vertical grid
level, $V_1$ is the resolved wind speed at the first grid level and
$\average{\bullet}$ represents the horizontal planar average. The mean
sub-filter stresses are computed using Monin-Obukhov similarity
law~\cite{Etling1996}. The formulation for the potential temperature flux could
potentially experience a singularity when the mean resolved potential
temperature at the first grid level $\average{\widetilde{\theta}_1}$ crosses the
reference temperature $\theta_{\circ}$. Hence the model for the sub-filter
potential temperature flux is reformulated using the Monin-Obukhov similarity
law as
\begin{equation}
(\tau_{\theta z})_0 = \left ( \frac{u_{\tau} \kappa}{ \phi_h} \right ) \frac{V_1 ( \average{\widetilde{\theta}_1} - \theta_{\circ} ) + \average{V_1} (\widetilde{\theta}_1 - \average{\widetilde{\theta}_1} ) }{ \average{V_1} },
\end{equation}
where $\phi_h$ is the non-dimensional potential temperature gradient from
Monin-Obukhov similarity law~\cite{Etling1996}.

@@

@@

\subsection{\label{sec:CFDsetup}Computational setup}

### 2020-12-04 Paper/results.tex
Added nomenclature and fixed notationi

%%%%%%%%%%%%%%% GRID STUDY: INTEGRAL LENGTH %%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
\caption{\label{tab:GridStudySetup} The setup for the grid resolution study} \centering
\begin{tabular}{ccccc}
  \hline
  Case              & dx [m] & dy [m] & dz [m] & Domain size \\
  \hline
  Nalu-Wind coarse  &  10.0  & 10.0   & 2.5 & 1.5 km $\times$ 1.5km $\times$ 1.0 km  \\
  Nalu-Wind medium  &   5.0  &  5.0   & 2.5 & 1.5 km $\times$ 1.5 km $\times$ 1.0 km  \\
  Nalu-Wind fine    &   2.5  &  2.5   & 2.5 & 1.0 km $\times$ 1.0 km $\times$ 1.0 km  \\
% AMR-wind fine     &   2.5  &  2.5   & 2.5 & 0.0 \\
\hline
\end{tabular}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{equation}
  f_{Ny} = \frac{1}{2\times \textrm{dt}}
\end{equation}
Secondly, the maximum resolvable frequency of convecting fine scale
eddies on a mesh with grid size $\Delta$ and average horizontal wind
speed $\bar{U}_{horiz}$ can be estimated as
\begin{equation}
  \label{eq:fmax}
  f_{max} = \frac{0.6\bar{U}_{horiz}}{N\Delta}.
\end{equation}
Here, Eq. \ref{eq:fmax} assumes that the turbulent eddies convect
with a velocity of $0.6\bar{U}_{horiz}$ and $N=8$ is chosen for the
purposes of this study.  Due to the alignment of the flow with the
mesh directions, the grid size is defined as $\Delta = \sqrt{2}\cdot \textrm{dx}$.
In all cases considered here, $f_{max} < f_{Ny}$, so the limiting
constraint for resolving the high frequency components was the mesh
resolution.

@@

%%%%%%%%%%%%%%% GRID STUDY: INTEGRAL LENGTH %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table} %[h!]
\caption{\label{tab:GridStudyLscale} The calculated turbulent integral
  lengthscale for each of the grid resolutions} \centering
\begin{tabular}{ccccc}
  \hline
  Case              & dx [m] & Longitudinal L  & Lateral L \\
  \hline
  Nalu-Wind coarse  &  10.0  & 36.4 m         & 0.0 m     \\
  Nalu-Wind medium  &   5.0  & 21.5 m         & 4.52 m    \\
  Nalu-Wind fine    &   2.5  & 25.3 m         & 5.92 m    \\
\hline
\end{tabular}
\end{table}

>>>>

%%%%%%%%%%%%%%% GRID STUDY: INTEGRAL LENGTH %%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
\caption{\label{tab:GridStudySetup} The setup for the grid resolution study} \centering
\begin{tabular}{ccccc}
  \hline
  Case              & $\Delta_x$ & $\Delta_y$ & $\Delta_z$ & Domain size \\
  \hline
  Nalu-Wind coarse  &  10.0 m  & 10.0 m & 2.5 m & 1.5 km $\times$ 1.5km $\times$ 1.0 km  \\
  Nalu-Wind medium  &   5.0 m &  5.0 m  & 2.5 m & 1.5 km $\times$ 1.5 km $\times$ 1.0 km  \\
  Nalu-Wind fine    &   2.5 m &  2.5 m  & 2.5 m & 1.0 km $\times$ 1.0 km $\times$ 1.0 km  \\
% AMR-wind fine     &   2.5  &  2.5   & 2.5 & 0.0 \\
\hline
\end{tabular}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{equation}
  f_{Ny} = \frac{1}{2\times \textrm{dt}}
\end{equation}
Secondly, the maximum resolvable frequency of convecting fine scale
eddies on a mesh with grid size $\Delta$ and average horizontal wind
speed $\bar{U}_{horiz}$ can be estimated as
\begin{equation}
  \label{eq:fmax}
  f_{max} = \frac{0.6\bar{U}_{horiz}}{N\Delta}.
\end{equation}
Here, Eq. \ref{eq:fmax} assumes that the turbulent eddies convect
with a velocity of $0.6\bar{U}_{horiz}$ and $N=8$ is chosen for the
purposes of this study.  Due to the alignment of the flow with the
mesh directions, the grid size is defined as $\Delta = \sqrt{2} \Delta_x $.
In all cases considered here, $f_{max} < f_{Ny}$, so the limiting
constraint for resolving the high frequency components was the mesh
resolution.

@@

%%%%%%%%%%%%%%% GRID STUDY: INTEGRAL LENGTH %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table} %[h!]
\caption{\label{tab:GridStudyLscale} The calculated turbulent integral
  lengthscale for each of the grid resolutions} \centering
\begin{tabular}{ccccc}
  \hline
  Case              & $\Delta_x$ & Longitudinal L  & Lateral L \\
  \hline
  Nalu-Wind coarse  &  10.0 m  & 36.4 m         & 0.0 m     \\
  Nalu-Wind medium  &   5.0 m & 21.5 m         & 4.52 m    \\
  Nalu-Wind fine    &   2.5 m & 25.3 m         & 5.92 m    \\
\hline
\end{tabular}
\end{table}
